---
title: "AREE RNA-seq Workflow Report"
subtitle: "`r params$study_id` / `r params$comparison_id`"
format:
  html:
    embed-resources: true
    toc: true
    theme: cosmo
params:
  study_id: "UNKNOWN_STUDY"
  comparison_id: "UNKNOWN_COMPARISON"
  mode: "processed_results_harmonization"
  standardized_tsv: "standardized.tsv"
  manifest_json: "manifest.json"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(jsonlite)
library(ggplot2)
`%||%` <- function(a, b) if (is.null(a)) b else a
```

## Run summary

This report was generated by the AREE `workflows/rnaseq` Nextflow scaffold.

```{r read-manifest}
manifest <- tryCatch(
  fromJSON(params$manifest_json, simplifyVector = TRUE),
  error = function(e) NULL
)

if (!is.null(manifest)) {
  summary_df <- data.frame(
    field = c("workflow_name", "workflow_version", "mode", "study_id",
              "comparison_id", "workflow_start"),
    value = c(
      manifest$workflow_name %||% NA,
      manifest$workflow_version %||% NA,
      manifest$mode %||% NA,
      manifest$study_id %||% NA,
      manifest$comparison_id %||% NA,
      manifest$workflow_start %||% NA
    )
  )
  knitr::kable(summary_df, col.names = c("Field", "Value"))
} else {
  cat("No manifest JSON could be read at:", params$manifest_json, "\n")
}
```

## Warnings

```{r warnings}
if (!is.null(manifest) && length(manifest$warnings) > 0) {
  for (w in manifest$warnings) cat("- ", w, "\n")
} else {
  cat("No warnings recorded.\n")
}
```

If `mode` is `processed_results_harmonization`, note that raw FASTQ QC
(FastQC/fastp/Salmon mapping rate) could not be independently verified —
see the `warnings` array above and `docs/design.md` section 7.

## Differential expression summary

```{r read-de}
de <- tryCatch(
  read.delim(params$standardized_tsv, sep = "\t", stringsAsFactors = FALSE),
  error = function(e) NULL
)

if (!is.null(de)) {
  n_sig <- sum(de$padj < 0.05, na.rm = TRUE)
  cat(sprintf(
    "Loaded %d features. %d features have padj < 0.05.\n",
    nrow(de), n_sig
  ))
  knitr::kable(head(de[order(de$padj, na.last = TRUE), ], 10))
} else {
  cat("No standardized DE table could be read at:", params$standardized_tsv, "\n")
}
```

## Volcano plot

```{r volcano, fig.width=7, fig.height=5}
if (!is.null(de)) {
  de$neg_log10_padj <- -log10(pmax(de$padj, 1e-300))
  de$sig <- ifelse(!is.na(de$padj) & de$padj < 0.05 & abs(de$log2FoldChange) > 1,
                    "significant", "not significant")

  ggplot(de, aes(x = log2FoldChange, y = neg_log10_padj, color = sig)) +
    geom_point(alpha = 0.7) +
    scale_color_manual(values = c("significant" = "firebrick", "not significant" = "grey70")) +
    geom_vline(xintercept = c(-1, 1), linetype = "dashed", color = "grey40") +
    geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "grey40") +
    labs(
      title = paste("Volcano plot:", params$study_id, params$comparison_id),
      x = "log2 Fold Change",
      y = expression(-log[10]("adjusted p-value")),
      color = NULL
    ) +
    theme_minimal()
} else {
  cat("Volcano plot skipped: no DE table available.\n")
}
```

## Provenance

Source file checksums, tool versions, and parameters are recorded in
`r params$manifest_json`, emitted by the `EMIT_MANIFEST` process. This
report does not duplicate that data verbatim; consult the manifest for the
authoritative, machine-readable record.

---

*Generated by the AREE `workflows/rnaseq` scaffold. This report template has
not been rendered against real pipeline outputs in this build — see
`workflows/rnaseq/README.md`.*
