---
title: "AREE Metabolomics Workflow Report"
subtitle: "`r params$study_id` / `r params$comparison_id`"
format:
  html:
    embed-resources: true
    toc: true
    theme: cosmo
params:
  study_id: "STUDY_ID"
  comparison_id: "COMPARISON_ID"
  mode: "processed_results_harmonization"
  standardized_tsv: "standardized.tsv"
  manifest_json: "manifest.json"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(jsonlite)
library(ggplot2)
library(dplyr)
library(knitr)

features <- read.delim(params$standardized_tsv, stringsAsFactors = FALSE)
manifest <- tryCatch(fromJSON(params$manifest_json), error = function(e) list())
```

## Run summary

This report was generated by the **AREE metabolomics workflow scaffold**
(`workflows/metabolomics/`). It is rendered from the standardized feature
table and the accompanying provenance manifest — it does not re-derive any
numbers, it only visualizes what those two files already contain.

- **Study:** `r params$study_id`
- **Comparison:** `r params$comparison_id`
- **Mode:** `r params$mode`
- **Date generated:** `r if (!is.null(manifest$date_generated)) manifest$date_generated else "unknown"`

```{r mode-warning, results='asis'}
if (identical(params$mode, "processed_results_harmonization")) {
  cat(
    "> **Note:** this comparison was ingested in `processed_results_harmonization` ",
    "mode. Upstream normalization and pooled-QC-sample CV metrics were **not** ",
    "independently verified by this workflow — they reflect the source study's ",
    "own pipeline, if reported at all. See `docs/design.md` section 7.\n"
  )
}
```

## Annotation confidence breakdown

Simplified MSI-style scale: 1 = confirmed by standard, 2 = putatively
annotated (spectral match), 3 = putative class, 4 = unknown.

```{r annotation-confidence}
conf_counts <- features |>
  count(annotation_confidence_level, name = "n_features") |>
  arrange(annotation_confidence_level)

kable(conf_counts, col.names = c("Annotation confidence level", "N features"))
```

```{r annotation-confidence-plot, fig.width=6, fig.height=3.5}
ggplot(conf_counts, aes(x = factor(annotation_confidence_level), y = n_features)) +
  geom_col(fill = "#3B7EA1") +
  labs(x = "Annotation confidence level (MSI-style, 1=best)", y = "Number of features") +
  theme_minimal()
```

## Metabolite class summary

```{r metabolite-class}
class_counts <- features |>
  count(metabolite_class, name = "n_features", sort = TRUE)

kable(class_counts, col.names = c("Metabolite class", "N features"))
```

## Normalization / QC metrics (upstream)

QC metrics captured from upstream intake/annotation/normalization steps (raw
mode) are embedded in the manifest under `qc_metrics.upstream_step_qc`; in
processed-results mode this section is typically empty (see note above).

```{r qc-metrics, results='asis'}
qc <- manifest$qc_metrics
if (!is.null(qc)) {
  cat(sprintf(
    "- N features: %s\n- N features at annotation confidence 1 or 2: %s (%s%%)\n",
    qc$n_features %||% "NA",
    qc$n_annotation_confidence_level_1_or_2 %||% "NA",
    qc$pct_annotation_confidence_level_1_or_2 %||% "NA"
  ))
} else {
  cat("No qc_metrics found in manifest.\n")
}
```

## Volcano plot

```{r volcano, fig.width=6.5, fig.height=5}
plot_df <- features |>
  mutate(
    neg_log10_p = -log10(pmax(pvalue, 1e-300)),
    sig = case_when(
      is.na(padj) ~ "NA",
      padj < 0.05 & log2FC > 0 ~ "up (padj<0.05)",
      padj < 0.05 & log2FC < 0 ~ "down (padj<0.05)",
      TRUE ~ "not significant"
    )
  )

ggplot(plot_df, aes(x = log2FC, y = neg_log10_p, color = sig)) +
  geom_point(alpha = 0.75) +
  scale_color_manual(values = c(
    "up (padj<0.05)" = "#B2182B",
    "down (padj<0.05)" = "#2166AC",
    "not significant" = "grey70",
    "NA" = "grey85"
  )) +
  labs(x = "log2 fold change", y = expression(-log[10](p-value)), color = NULL) +
  theme_minimal()
```

## Top differential features

```{r top-features}
top <- features |>
  arrange(padj, pvalue) |>
  head(20) |>
  select(feature_id, putative_metabolite_name, annotation_confidence_level,
         metabolite_class, log2FC, pvalue, padj)

kable(top)
```

## Provenance

```{r provenance, results='asis'}
if (length(manifest) > 0) {
  cat(sprintf("- Workflow: %s (%s)\n", manifest$workflow_name %||% "NA", manifest$workflow_version %||% "NA"))
  cat(sprintf("- Generated by: %s\n", manifest$generated_by %||% "NA"))
  if (length(manifest$warnings) > 0) {
    cat("- Warnings:\n")
    for (w in manifest$warnings) cat(sprintf("  - %s\n", w))
  }
} else {
  cat("No manifest.json found alongside this report.\n")
}
```

---

*This report and the workflow that produced it are part of the AREE
(Aquaculture Resilience Evidence Engine) metabolomics scaffold. It is an
unexecuted structural template in this build — see
`workflows/metabolomics/README.md` for what has and has not been run.*
