---
title: "AREE Methylation Workflow Report"
subtitle: "`r params$study_id` / `r params$comparison_id`"
format:
  html:
    embed-resources: true
    toc: true
    theme: cosmo
params:
  study_id: "UNKNOWN_STUDY"
  comparison_id: "UNKNOWN_COMPARISON"
  standardized_table: "dmr_standardized.tsv"
  manifest_json: "manifest.json"
  mode: "processed_results_harmonization"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(jsonlite)
```

## Scaffold status

**This report template has not been executed against real data in this
build.** It is a structurally complete Quarto document designed to render
against the outputs of `workflows/methylation/main.nf`. When run against the
shipped demo processed-results table
(`data/demo/methylation/GIGAS_PATH03_OsHV1_challenge_vs_sham_dmr_demo.tsv`)
it would produce real summary tables and plots from that (clearly-labeled
simulated) data; it has not been rendered as part of this task.

```{r load-data}
dmr <- read.delim(params$standardized_table, sep = "\t", stringsAsFactors = FALSE)
manifest <- tryCatch(fromJSON(params$manifest_json), error = function(e) NULL)
```

## Run identification

- **Study ID:** `r params$study_id`
- **Comparison ID:** `r params$comparison_id`
- **Mode:** `r params$mode`
- **Regions in standardized table:** `r nrow(dmr)`

```{r manifest-summary}
if (!is.null(manifest)) {
    cat("Workflow:", manifest$workflow_name, manifest$workflow_version, "\n")
    cat("Generated:", manifest$date_generated, "\n")
    if (length(manifest$warnings) > 0) {
        cat("\nWarnings:\n")
        for (w in manifest$warnings) cat(" -", w, "\n")
    }
} else {
    cat("No manifest.json found alongside this table; provenance fields unavailable.")
}
```

## Hyper vs. hypo methylation

```{r direction-table}
knitr::kable(table(Direction = dmr$direction), col.names = c("Direction", "N regions"))
```

```{r direction-plot, fig.width=5, fig.height=4}
if (nrow(dmr) > 0) {
    counts <- table(dmr$direction)
    barplot(
        counts,
        col = c(hyper = "#c0392b", hypo = "#2980b9")[names(counts)],
        main = "DMRs by direction",
        ylab = "Number of regions"
    )
} else {
    plot.new()
    title("No regions to plot")
}
```

## Genomic context distribution

```{r context-table}
knitr::kable(table(Context = dmr$annotation_context), col.names = c("Annotation context", "N regions"))
```

```{r context-plot, fig.width=6, fig.height=4}
if (nrow(dmr) > 0 && "annotation_context" %in% names(dmr)) {
    ctx_counts <- sort(table(dmr$annotation_context), decreasing = TRUE)
    barplot(
        ctx_counts,
        las = 2,
        main = "Regions by genomic context",
        ylab = "Number of regions"
    )
} else {
    plot.new()
    title("No annotation context data to plot")
}
```

## Intergenic / gene-less regions

Regions with no associated `gene_id` are intentionally retained in the
standardized table (not dropped) and are harmonized downstream as
`genomic_region` features with `mapping_confidence = unresolved`
(see `src/harmonize/methylation.py`).

```{r intergenic-summary}
n_no_gene <- sum(is.na(dmr$gene_id) | trimws(as.character(dmr$gene_id)) == "")
cat(n_no_gene, "of", nrow(dmr), "regions have no gene_id (kept as intergenic/unresolved).\n")
```

## Full standardized region table

```{r full-table}
knitr::kable(dmr)
```

## Provenance

```{r manifest-raw}
if (!is.null(manifest)) {
    str(manifest, max.level = 1)
} else {
    cat("No manifest available.")
}
```
