---
title: "AREE Proteomics Workflow Report"
subtitle: "`r params$study_id` / `r params$comparison_id`"
format:
  html:
    embed-resources: true
    toc: true
params:
  standardized_table: "standardized.tsv"
  manifest_json: "manifest.json"
  study_id: "STUDY_ID"
  comparison_id: "COMPARISON_ID"
---

<!--
STATUS: real Quarto/R Markdown template with executable R chunks (ggplot2
volcano plot, missingness summary, top-hits table). It has NOT been rendered
against a real run's outputs in this build -- see
workflows/proteomics/README.md. It is written to run unmodified against the
files produced by modules/proteomics/standardize_output.nf and
modules/proteomics/emit_manifest.nf, whether those came from raw_reanalysis
or processed_results_harmonization mode.
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(jsonlite)
library(dplyr)
library(ggplot2)

standardized <- read.delim(params$standardized_table, sep = "\t", stringsAsFactors = FALSE)
manifest <- fromJSON(params$manifest_json)
```

## Run summary

- **Study**: `r params$study_id`
- **Comparison**: `r params$comparison_id`
- **Mode**: `r manifest$mode`
- **Workflow version**: `r manifest$workflow_version`
- **Date generated**: `r manifest$date_generated`
- **Provenance level**: `r manifest$provenance_level`

```{r warnings-block, results='asis'}
if (length(manifest$warnings) > 0) {
  cat("### Warnings\n\n")
  for (w in manifest$warnings) {
    cat(sprintf("- %s\n", w))
  }
} else {
  cat("No warnings recorded for this run.\n")
}
```

## QC metrics

```{r qc-table}
qc <- as.data.frame(manifest$qc_metrics)
knitr::kable(qc)
```

## Missingness distribution

Per-protein `missingness_percent` from the standardized evidence-ready table.
Values of `NA` indicate the source study/mode did not report missingness
(this is left explicit, not imputed as zero, per docs/design.md section 4).

```{r missingness-plot}
if (all(is.na(standardized$missingness_percent))) {
  cat("missingness_percent is entirely NA for this comparison -- no plot rendered.\n")
} else {
  ggplot(standardized, aes(x = missingness_percent)) +
    geom_histogram(binwidth = 5, boundary = 0, fill = "#4C72B0", color = "white") +
    labs(
      x = "Missingness (%) per protein",
      y = "Number of proteins",
      title = "Per-protein missingness"
    ) +
    theme_minimal()
}
```

## Volcano plot

```{r volcano-plot}
plot_df <- standardized %>%
  filter(!is.na(log2FC), !is.na(padj)) %>%
  mutate(
    neg_log10_padj = -log10(pmax(padj, 1e-300)),
    significant = padj < 0.05 & abs(log2FC) >= 1
  )

if (nrow(plot_df) == 0) {
  cat("No rows with non-missing log2FC/padj -- no volcano plot rendered.\n")
} else {
  ggplot(plot_df, aes(x = log2FC, y = neg_log10_padj, color = significant)) +
    geom_point(alpha = 0.7) +
    geom_vline(xintercept = c(-1, 1), linetype = "dashed", color = "grey50") +
    geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "grey50") +
    scale_color_manual(values = c(`TRUE` = "#C44E52", `FALSE` = "grey60")) +
    labs(
      x = "log2 fold change",
      y = expression(-log[10](adjusted~p)),
      title = "Differential protein abundance",
      color = "padj < 0.05 & |log2FC| >= 1"
    ) +
    theme_minimal()
}
```

## Top differentially abundant proteins

```{r top-hits}
top_hits <- standardized %>%
  filter(!is.na(padj)) %>%
  arrange(padj) %>%
  head(20)

knitr::kable(top_hits)
```

## Reproducibility

```{r manifest-dump}
str(manifest, max.level = 1)
```

This report was generated from `r params$standardized_table` and
`r params$manifest_json` by the AREE proteomics workflow scaffold. It is
intended to accompany, not replace, the standardized TSV as the
machine-readable artifact consumed by `aree harmonize`.
